---
title: "Interpolated Autocorrelation Analysis"
format: html
---


For the visualizer and the annotator, the satellite images were obtained from Ersi World Imagery.
Images were kept saved for offline use.
Leaflet was used to make the interactive labler.



## CSV Columns Documentation

The data comes from `../04_lane/gpsErrors.csv`, made by the lane interpolation code in 04.
Each row is an original camera GPS reading compared with the modeled car position at the same time.
Only the car position is interpolated, not the GPS readings.
The main reading ID is `camera` plus `readingIndex`, which starts at 1 for each camera and continues across clips.
`clip` and `sampleIndex` are kept as source references; autocorrelation still uses `timeSec`, not the reading index.

| Column name | Description | Data type in R | Example(s) | Small set / range |
|---|---|---|---|---|
| `camera` | Which camera the reading came from. | `character` (text) | `garmin`, `gopro` | `garmin`, `gopro`, `redtiger`, `rove` |
| `readingIndex` | Chronological GPS reading number for this camera across the drive, not a video frame number. | `integer` | `1`, `2`, `3` | `1` to this camera's total number of readings |
| `clip` | Original MP4 clip, kept to trace the reading to its source. | `character` (text) | `GRMN0213.MP4`, `GX010003.MP4` | — |
| `sampleIndex` | Original zero-based record index within the clip, kept as a source reference. | `integer` | `33`, `3082` | — |
| `timeSec` | Seconds since July 13th 2026, 5:28AM CDT. | `numeric` (decimal) | `0`, `0.1`, `0.2` | — |
| `latitude` | Camera GPS latitude, in decimal degrees. | `numeric` (decimal) | `40.8110815` | — |
| `longitude` | Camera GPS longitude, in decimal degrees. | `numeric` (decimal) | `-96.6894486` | — |
| `speedMps` | Camera-reported speed, converted to meters per second. | `numeric` (decimal) | `0`, `0.022` | — |
| `fix` | Whether the GPS fix was accepted; GoPro 2D and 3D fixes are accepted. | `integer` | `1` | `0` = not accepted; `1` = accepted. This CSV only contains `1`. |
| `segment` | Index of the interval between annotation points. | `integer` | `1`, `2` | `1`–`374` in this drive |
| `referenceLatitude` | Interpolated car latitude, in decimal degrees. | `numeric` (decimal) | `40.8111466` | — |
| `referenceLongitude` | Interpolated car longitude, in decimal degrees. | `numeric` (decimal) | `-96.6894028` | — |
| `distanceM` | Distance between the interpolated car and GPS point, in meters. | `numeric` (decimal) | `8.197385`, `6.730981` | — |
| `directionRad` | Direction from the interpolated car to the GPS point, in radians, counterclockwise from grid east. | `numeric` (decimal) | `-2.035405`, `2.642391` | −π to π (about −3.14159 to 3.14159) |



Examples are taken from the CSV and rounded where needed. A dash means no short value list or range is shown.
Missing values are blank in the CSV and read as `NA` in R; these are separate from the values listed above.


The quality flags are now in `../../audit/referenceIntervals.csv`, matched using "segment".
"referenceOk" is TRUE if the interval passed the model quality checks; "reason" explains any flags.
The default `preserveV4Exclusions` setting in `../../settings.json` keeps the same exclusions as the original analysis.
Set it to FALSE to include all modeled intervals with accepted GPS fixes instead.

More info on the flags:
path_obd_distance_disagreement : modeled path distance divided by integrated OBD distance falls outside of (0.75,1.25)
anchor_confidence_below_60: at least one endpoint annotation had a confidence score below 60%
construction_or_occluded_lane_uncertainty: interval belonged to route section previously flagged for uncertaihn lane placement.
flagged_v4_geometry: the saved v4 geometry retained an unsuported/conflict flag
elevation_grade_needs_review: The elevation profile implied a slope above 15% grade. possibly not accurate.
conflicting_stop_coordinates: points 0-1 had this, stop with different positions, but this was not severe.
coincident_anchors_without_held_stop: points 118-119 shared a position without a complete held-stop label



```{r startup}
library(ggplot2)
source("../../buildChecks.R")
assertCurrentBuild("../..")

gps <- read.csv("../04_lane/gpsErrors.csv", na.strings = c("", "NA"))
intervals <- read.csv("../../audit/referenceIntervals.csv", na.strings = c("", "NA"))
settings <- jsonlite::fromJSON("../../settings.json")
cameras <- c("garmin", "gopro", "redtiger", "rove")

```


Get the error for each gps, in terms of distance and direction in radians.
The new CSV already has these calculated using `sqrt(east^2 + north^2)` and `atan2(north, east)` in projected meters.
Keep excluded readings as NA so time gaps are not removed.


```{r error dist dir}
gps$radius <- gps$distanceM
gps$theta <- gps$directionRad
gps$usable <- gps$fix == 1 & is.finite(gps$radius)
if (settings$preserveV4Exclusions) {
  gps$usable <- gps$usable & intervals$referenceOk[match(gps$segment, intervals$segment)]
}
gps$usable[is.na(gps$usable)] <- FALSE
gps$radius[!gps$usable] <- NA
gps$theta[!gps$usable | (!is.na(gps$radius) & gps$radius == 0)] <- NA
```


Calculate autcorrelation


```{r get autocor}
results <- list()
for (camera in cameras) {
  e <- gps[gps$camera == camera, ]
  step <- if (camera == "gopro") 0.1 else 1
  slot <- round(e$timeSec / step)
  stopifnot(!anyDuplicated(slot), min(slot) >= 0,
            max(abs(e$timeSec - slot * step)) < 0.002)

  radius <- theta <- rep(NA_real_, max(slot) + 1)
  radius[slot + 1] <- e$radius
  theta[slot + 1] <- e$theta
  stopifnot(sum(is.finite(radius)) > 1, var(radius, na.rm = TRUE) > 0)

  distance_acf <- acf(radius, lag.max = 300 / step,
                      plot = FALSE, na.action = na.pass)
  direction_cov <- acf(cbind(cos(theta), sin(theta)), type = "covariance",
                       lag.max = 300 / step, plot = FALSE,
                       na.action = na.pass)$acf
  direction_acf <- direction_cov[, 1, 1] + direction_cov[, 2, 2]
  stopifnot(is.finite(direction_acf[1]), direction_acf[1] > 1e-20)
  direction_acf <- direction_acf / direction_acf[1]

  lag_seconds <- as.numeric(distance_acf$lag) * step
  results[[camera]] <- data.frame(
    camera, lag_seconds = rep(lag_seconds, 2),
    component = rep(c("Distance", "Direction"), each = length(lag_seconds)),
    autocorrelation = c(as.numeric(distance_acf$acf), direction_acf))
}
result <- do.call(rbind, results)
```



```{r plot}
#| fig-width: 10
#| fig-height: 4.5
#| fig-alt: "Static distance and direction autocorrelation versus lag in seconds for four receivers, relative to the interpolated v4 lane-center car."
result$camera <- factor(result$camera, levels = cameras,
                        labels = c("Garmin", "GoPro", "REDTIGER", "ROVE"))
result$component <- factor(result$component, levels = c("Distance", "Direction"))
plot <- ggplot(result, aes(lag_seconds, autocorrelation, color = camera)) +
  geom_line(linewidth = 0.7) +
  facet_wrap(~ component, nrow = 1) +
  scale_color_manual(values = c("#7C3AED", "#0284C7", "#D97706", "#15803D")) +
  scale_x_continuous(breaks = seq(0, 300, 60)) +
  labs(x = "Lag (seconds)", y = "Autocorrelation", color = NULL) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom", panel.grid.minor = element_blank())

ggsave("interpolated_autocorrelation.png", plot, width = 10, height = 4.5, dpi = 300)
plot
```
